---
title: "MM_Branch -- Affordable Housing Project"
output: html_document
date: "2022-11-23"
---

```{r}
library(ggplot2)
library(tidyr)
library(tidyverse)
library(dplyr)
library(readr)
library(sf)
library(stringr)
library(data.table)
```


```{r}
assessments <- read_csv("Data/assessments.csv")
barnes <- read_csv("Data/barnes.csv")
filtered_sales <- read_csv("Data/filtered_sales.csv")
full_sales <- read_csv("Data/full_sales.csv")
LIHTC <- read_csv("Data/LIHTC_updated.csv")
property_details <- read_csv("Data/property_details.csv")
```
#### Filter LIHTC to Developements in service since 2000
#### Remove invalid years from YR_PIS column
#### Filter to include only valid columns

```{r}
LIHTC_2000 <- LIHTC %>% 
  filter(YR_PIS >= 2000) %>% 
  filter(YR_PIS != 8888, YR_PIS != 9999)
LIHTC_2000_mod <- LIHTC_2000 %>% 
  select(HUD_ID, PROJECT, YR_PIS, LATITUDE, LONGITUDE)
```
#### Filter Barnes Data to include 5 columns: HUD_ID, YR_PIS, lat, long, Project (name)
#### Convert YR_PIS to numeric column
#### Rename Columns to associated LIHTC column names
```{r}
barnes$'Affordability Term Start Date' <- str_sub(barnes$'Affordability Term Start Date', start=-4, end=-1) %>%
  as.numeric()
barnes_mod <- barnes[, c('HUD_ID', 'Development Name', 'Affordability Term Start Date', 'lat', 'lng')]
colnames(barnes_mod) <- c('HUD_ID', 'PROJECT', 'YR_PIS', 'LATITUDE', 'LONGITUDE')
```
#### Join LIHTC_2000 and Barnes Data -- name = "developements"

```{r}
developments <- rbind(LIHTC_2000_mod, barnes_mod)
```

#### Separate 'centroid' lat and long values in property_details
#### Convert property_details to sf
```{r}
property_details$latlong <- stringr::str_extract(string = property_details$centroid, pattern = "(?<=\\().*(?=\\))")
property_details[c('long', 'lat')] <- str_split_fixed(property_details$latlong,',', 2)
pd_sf <- st_as_sf(property_details, coords = c("long", "lat"), crs = 4326)
```

#### Convert developments into sf 
```{r}
developments_sf <- st_as_sf(developments, coords = c("LONGITUDE", "LATITUDE"), crs = 4326)
```

#### Get Nearest Development to Property -- assign to 'nearest'
#### Get distance between developement and property -- assign to 'dist'
#### bind the two together, including only HUD_ID, YR_PIS, Project
#### add the distance column to get_nearest sf
#### mutate to convert distance column to a numeric value
```{r}
nearest <- st_nearest_feature(pd_sf, developments_sf)
dist <- st_distance(pd_sf, developments_sf[nearest,], by_element = TRUE) %>%
    units::set_units(mi)
get_nearest <- cbind(pd_sf, st_drop_geometry(developments_sf[nearest, c("HUD_ID", "YR_PIS", "PROJECT")]))
get_nearest$dist = dist
get_nearest <- get_nearest %>% 
  mutate(dist = as.numeric(dist))
```

#### Filter the homes down to those that are within one mile of an affordable housing development.
```{r}
pd_within_mi <- get_nearest %>% 
  filter(dist <= 1)
```

#### Combine filtered_sales to pd_within_mi to include sales price and year
#### Keep necessary columns from combined sf dataset
#### Create new column, called 'difference' to account for the difference in property purchase and the nearest developement year placed in service
```{r}
dropdup_fs <- filtered_sales %>%
  distinct()
pd_within_mi_sales <- inner_join(pd_within_mi, dropdup_fs, by = "apn", copy = FALSE, suffix = c(".pd_within_mi", ".dropdup_fs"))
pd_within_mi_sales <- pd_within_mi_sales %>% 
  select(apn, tract, year_built, amount, ownerdate, land_area, square_footage, story_height, building_condition, number_of_rooms, number_of_beds, number_of_baths, number_of_half_bath, number_of_fixtures, PROJECT, HUD_ID, dist, YR_PIS, geometry)
pd_within_mi_sales$ownerdate <- pd_within_mi_sales$ownerdate %>% 
  format(format = "%Y") %>% 
  as.numeric()
pd_within_mi_sales <- pd_within_mi_sales %>% 
  mutate(difference = (YR_PIS - ownerdate))
```

#### calculate a new column called "group"
```{r}
pd_within_mi_sales <- pd_within_mi_sales %>%
  mutate(group = 
           case_when(
             dist < 0.5 & difference >= 2 & difference <= 5 ~ "pre",
             dist < 0.5 & difference >= 0 & difference < 2 ~ "mid",
             dist < 0.5 & difference < 0 ~ "post",
             dist > 0.5 & difference >= 0 & difference <= 5 ~ "outside",
             TRUE ~ "other"
    )
  )
```

#### Filter out all rows whose group is "other"
```{r}
pd_within_mi_sales <- pd_within_mi_sales %>% 
  filter(group != "other")
```

#### Create a column "Tpost" that, for homes in the "post" group gives the number of years that the sale occurred after the housing development was placed in service
```{r}
pd_within_mi_sales <- pd_within_mi_sales %>%
    mutate(Tpost = if_else(group == "post", (ownerdate - YR_PIS), 0))
```

#### Create a column for "age" of home when sold
```{r}
pd_within_mi_sales <- pd_within_mi_sales %>% 
  mutate(age = (ownerdate - year_built))

```

#### Filter to any property sold 5 years before development to 5 years after
```{r}
prop_sold_5yrs <- pd_within_mi_sales %>% 
  filter(difference >= -5 & difference <= 5)
```

#### Filter any properties that sold prior to they were built (~3k properties)
```{r}
prop_sold_5yrs <- prop_sold_5yrs %>% 
  filter(age >= 0)
```

#### Convert land_areas into numeric and tract into character for better model fitting
```{r}
prop_sold_5yrs[c("land_area_acres", "acres")] <- str_split_fixed(prop_sold_5yrs$land_area,' ', 2)

prop_sold_5yrs <- subset(prop_sold_5yrs, select = -c(acres) )
prop_sold_5yrs$land_area_acres <- prop_sold_5yrs$land_area_acres%>%
  as.numeric()

prop_sold_5yrs$tract <- prop_sold_5yrs$tract%>%
  as.character()
```

#### Then build a linear model with target variable the sales amount using property features
```{r}
simple_regression_model <- lm(amount ~ square_footage +  age + group + ownerdate + tract, data = prop_sold_5yrs)

summary(simple_regression_model)
```


#### Create a model with target being the log of the sale price
```{r}
simple_regression_model_log <- lm(log(amount) ~ square_footage +  age + group + ownerdate + tract, data = prop_sold_5yrs)

summary(simple_regression_model_log)
```

```{r}

```




